---
title: "HMC Diagnostics"
output: rmarkdown::html_vignette
---

This notebook runs a few statistical diagnostics on the logged variables of an HMC run. The experiment is defined through the  `path` variable which defines the directory where the logging was made. This notebook assumes that the files provided are:

- `log_gamma.npy`
- `log_lambda.npy`
- `parameter*.npy`
```{r}
library(reticulate)
np <- import("numpy")
library(coda)
library(ggplot2)

params <- "../etc/diagnostic/flux_nn/full_parameters.npy"
params <- np$load(params)
all_chains <- apply(apply(params, c(2, 3), function(col) {mcmc(col)}, simplify = FALSE), 1, function(col) {mcmc.list(col)}, simplify = FALSE)
```
# Geweke Z-score

We first look at the Geweke Z-score for all our variables. This test compares the distribution between the beginning and the end of each Markov chain.
```{r}
get.all.geweke <- function(all_chains){
  get.geweke <- function(chains){
    z <- geweke.diag(chains)
    z <- unlist(z, use.names = FALSE)
    z <- z[seq(1, length(z), 3)]
    return(z)
  }
  z <- sapply(all_chains, get.geweke)
  return(z)
}

z <- get.all.geweke(all_chains)
off <- sum(abs(z) > 2)
cat(sprintf("Out of %dx%d parameters, %d fail the Geweke Z-Test (%.1f%%)", dim(z)[1], dim(z)[2], off, off/length(z)*100))
```


# Gelman-Rubin test

The Gelman-Rubin test uses several Markov chains for each variable. It computes a ratio using both the within-chain variance and the between-chain variance. If the chain converges correctly, it should be close to 1.
```{r}
g <- sapply(all_chains, function(chains) {x <- gelman.diag(chains); x$psrf})
conf <- 1.5
off <- sum(g[2,] > conf)
cat(sprintf("With upper confidence interval limit at %.2f:\n\tOut of %d parameters, %d fail the Gelman-Rubin (%.1f%%)",conf, length(g[2,]), off, off/length(g[2,])*100))

conf <- 3
off <- sum(g[2,] > conf)
cat(sprintf("With upper confidence interval limit at %.2f:\n\tOut of %d parameters, %d fail the Gelman-Rubin (%.1f%%)",conf, length(g[2,]), off, off/length(g[2,])*100))
```
# Auto-correlation

Finally, for a few parameters at random, let's plot their auto-correlation for different lag values:

```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```

```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```

```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```

```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```
```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```
```{r}
idx <- sample(1:length(all_chains),1)
y <- all_chains[[idx]][[1]]
autocorr.plot(y, length(y)/5)
```
